Exact Asymptotics for Causal Mediation Analysis

### 2024-05-06 notes/William Analysis.Rmd
Re-name write-up
Implement binary, binary, fixed in R. SEs are currently wrong

```{r, boilerplate, eval=FALSE}
# Note: This code chunk is not run here, but it is called and run at the start of every subsequent section. Any changes made here will be duplicated throughout the rest of the document.

@@

>>>>

```{r, boilerplate, include=FALSE, cached=TRUE}
# Note: This code chunk is not run here, but it is called and run at the start of every subsequent section. Any changes made here will be duplicated throughout the rest of the document.

@@

# Binary Response, Binary Mediator, Fixed-Effects

For our last fixed-effect model, we now take both the response and mediator to be binary. I wrote some helper functions to compute relevant derivatives, all of which validated against a simple finite-difference approximation. See `Helpers.R`.

## Generate Data

As in the previous case, we set intercepts so that the means of both linear predictors are at least approximately zero. Here however, no residuals are required.

```{r}
a_0 = 0
b_0 = -0.5
```

We first demonstrate our analysis on a single dataset, then put this analysis inside a `for` loop. The first step is to set the sample size and generate some data. We use $\eta$ to denote the linear predictor for $M$, and $\zeta$ for the linear predictor of $Y$ without the $M$ term. That is, $\zeta = b_0 + b_2 X + W B_3$.

# Generate Y
zeta_vec = b_0 + b_1*M + b_2*X + W%*%B_3
p_Y_vec = expit(zeta_vec)
Y = rbinom(n, size = 1, prob = p_Y_vec)
```

We now fit logistic regression models to predict $M$ and $Y$, and extract output that we will need later.

```{r}
M_data = data.frame(M, X, W1 = W[, 1], W2 = W[, 2], W3 = W[,3])
M_model = glm(M ~ X + W1 + W2 + W3, data = M_data, family = binomial(link = "logit"))

a_hat = summary(M_model)$coefficients[, 1]
a_SE = summary(M_model)$coefficients[, 2]
a_cov = vcov(M_model)

Y_data = data.frame(Y, M, X, W1 = W[, 1], W2 = W[, 2], W3 = W[,3])
Y_model = glm(Y ~ M + X + W1 + W2 + W3, data = Y_data, family = binomial(link = "logit"))

b_hat = summary(Y_model)$coefficients[, 1]
b_SE = summary(Y_model)$coefficients[, 2]
b_cov = vcov(Y_model)
```

Next, we estimate the mediation effect. Note that, as in the contiuous outcome, binary mediator setting, our mediation effect depends on $X$ and $W$. We again use $X=0$ and $W=[1,1,1]$, although Samoilenko and Lefebvre (2023) use the sample means of these covariates.

# Linear predictors
eta_hat = a_0_hat + a_x_hat*x_pred + W_pred%*%A_2_hat
zeta_hat = b_0_hat + b_x_hat * x_pred + W_pred %*% B_3_hat

# Mediation effect
## See Helpers.R for the function get_odds_ratio
med_hat = get_odds_ratio(eta_hat, a_x, zeta_hat, b_x, b_m)
```

We now compute the gradient of the odds ratio and use the $\delta$-method to get the standard error.

```{r}
# Gradient of OR wrt regression coefficients
grad_med = d_OR_d_theta(eta_hat, a_x_hat, zeta_hat, b_x_hat, b_m_hat, x_pred, W_pred)

# Get asymptotic SE using delta method

## Build joint covariance matrix of regression coefficients
a_length = nrow(a_cov)
b_length = nrow(b_cov)
joint_cov = matrix(0, nrow = a_length + b_length, ncol = a_length +
b_length)
joint_cov[1:a_length, 1:a_length] = a_cov
joint_cov[(a_length + 1):(a_length + b_length), (a_length + 1):(a_length +
b_length)] = b_cov

## Convert to asymptotic covariance matrix
asymp_reg_cov = n * joint_cov

## Pre- and post-multiply asymptotic covariance by gradient of transformation
med_asymp_var = grad_med %*% asymp_reg_cov %*% grad_med

## Get small-sample standard error
med_asymp_SE = sqrt(med_asymp_var)
med_SE = med_asymp_SE / sqrt(n)
```

Putting everything together, we have an estimated mediation effect of `r med_hat`, with an estimated SE of `r med_SE`.

We now repeat the above analysis `r num_reps` times for each of various values of $n$. Values of $X$ and $W$ for which we compute the total effect are set at the beginning. We also re-run the boilerplate code (not shown).

```{r}
# Values of X and W for which we compute the total effect
x_pred = 0
W_pred = c(1,1,1)

# Values of a_0 and b_0
a_0 = 0
b_0 = -0.5
```

    ## Generate Y
    zeta_vec = b_0 + b_1*M + b_2*X + W%*%B_3
    p_Y_vec = expit(zeta_vec)
    Y = rbinom(n, size = 1, prob = p_Y_vec)

    ## Y
    Y_data = data.frame(Y, M, X, W1 = W[,1], W2 = W[,2], W3 = W[,3])
    Y_model = glm(Y ~ M + X + W1 + W2 + W3, data = Y_data, family = binomial(link = "logit"))

    ## Linear predictor for M
    eta_hat = a_0_hat + a_x_hat*x_pred + W_pred%*%A_2_hat
    zeta_hat = b_0_hat + b_x_hat * x_pred + W_pred %*% B_3_hat

    # Mediation effect
    ## See Helpers.R for the function get_odds_ratio
    med_hat = get_odds_ratio(eta_hat, a_x, zeta_hat, b_x, b_m)
    some_med_hats[[i]] = med_hat

    ## Compute gradient of mediation effect
    grad_med = d_OR_d_theta(eta_hat, a_x_hat, zeta_hat, b_x_hat, b_m_hat, x_pred, W_pred)

We now process the output from our simulation and summarize the results in a table. We give the mean and median of the estimated standard errors, as well as their relative errors (in $\%$) as estimates of the empirical standard error. See Table \ref{tab:results_bin_bin_fix}.

```{r results="asis", echo=FALSE}
## Print table
kbl(results, format = "latex", booktabs = T, caption = "Summary of SE estimates for mediation effect under continuous response, continuous mediator, fixed-effects model \\label{tab:results_bin_bin_fix}")
```
